---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

#### unzip the dataset
```{r echo=TRUE}
if(!dir.exists('./data')) {
  dir.create('./data')
}

if(!file.exists('data/activity.csv')) {
  unzip('activity.zip', exdir = './data')
}
```

#### Load the data into activity variable
```{r echo=TRUE}
activity <- read.csv('./data/activity.csv')
activityWoNA <- na.omit(activity)
```

## What is mean total number of steps taken per day?

Aggregate sum of steps per days and calculate mean and median
```{r echo=TRUE}
stepsPerDay <- aggregate(steps~date, data = activityWoNA, FUN = sum)
stepsPerDayMean <- mean(stepsPerDay$steps)
stepsPerDayMedian <-  median(stepsPerDay$steps)
```

Make a function to create the histogram
```{r echo=TRUE} 
library("ggplot2")
makeHistogram <- function(data) {
  stepsPerDayPlot <- ggplot(data = data, aes(x=steps))
  stepsPerDayPlot + geom_histogram(binwidth = 5000,colour="darkgreen", fill="lightgreen",alpha=0.4) +
                    ggtitle("Distribution of steps per day") + ylab("Frequency") + xlab("Total steps")
}
```

Histogram with ggplot2
``` {r echo=TRUE, fig.withd=10, fig.height=5, fig.align="center"}
makeHistogram(stepsPerDay)
```

The *mean* is `r mean(stepsPerDay$steps)` and the *Median* is `r median(stepsPerDay$steps)`

## What is the average daily activity pattern?

aggregate data by intervals
```{r echo=TRUE, fig.withd=10, fig.height=5, fig.align="center"}
meanStepsPerIntervals <- aggregate(steps~interval, data=activityWoNA, FUN = mean)
meanStepsIntervalsPlot <- ggplot(data=meanStepsPerIntervals, aes(x=interval,y=steps))
meanStepsIntervalsPlot+geom_line(color="#66cdaa")
intervalWithMax <- meanStepsPerIntervals[which.max(meanStepsPerIntervals$steps),]$interval
```

The **interval** with *max mean* steps is **`r intervalWithMax`**

## Imputing missing values

1. Get the total missing values
```{r echo=TRUE}
totalMissingValues <- activity$steps[is.na(activity$steps)]
```

There are **`r length(totalMissingValues)`** missing values.

2-3. Replace missing values with the mean in a new dataset
```{r echo=TRUE}
activityNARep <- merge(activity, meanStepsPerIntervals, by="interval")
activityNARep$steps.x[is.na(activityNARep$steps.x)] <- activityNARep$steps.y[is.na(activityNARep$steps.x)]
activityNARep$steps <- activityNARep$steps.x
activityNARep <- activityNARep[,c("interval", "date", "steps")]
```

4. Make a Histogram report
```{r echo=TRUE, fig.withd=10, fig.height=5, fig.align="center"}
stepsPerDayNARep <- aggregate(steps~date, data = activityNARep, FUN = sum)
makeHistogram(stepsPerDayNARep)
```

The *mean* is `r mean(stepsPerDayNARep$steps)` and the *Median* is `r median(stepsPerDayNARep$steps)`

## Are there differences in activity patterns between weekdays and weekends?

Add a column with label *weekday*  and *weekend* and aggregate the mean value per interval and the new column.
Produce a graph faceted by the new column (typeOfDay).
```{r echo=TRUE, fig.withd=10, fig.height=5, fig.align="center"}
activityNARep$typeOfDay <- ifelse(weekdays(as.Date(activityNARep$date,"%Y-%m-%d")) %in% c("dimanche", "samedi"), "weekend", "weekday")
stepsIntervalPerTypeOfDay <- aggregate(steps~interval+typeOfDay, activityNARep, FUN=mean)
weekDaysPlot <- ggplot(stepsIntervalPerTypeOfDay, aes(x=interval, y=steps, colour=factor(typeOfDay)))
weekDaysPlot + geom_line() + facet_wrap(~ typeOfDay, ncol=1)
```